---
title: "FML Assignment 3"
author: "Jenny Prashanth Buraga"
output: html_document
date: "2023-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First, Let us load the required Libraries and the dataset
```{r}
library(lattice)
library(ISLR)
library(ggplot2)
library(caret)
library(class)
library(e1071)
df <- read.csv("C:/Users/ssim/OneDrive/Desktop/Rhistory/UniversalBank.csv")
summary(df)
```

# Let us now partition the data into Training(60%) and Validation(40%) datasets
```{r}
set.seed(123)
index_train <- createDataPartition(df$CreditCard, p=0.6, list = F)
train_data <- df[index_train, ]
validation_data <- df[-index_train, ]
```
# Question A
# Create Pivot Table
```{r}
pivot_table <- xtabs(~ Online+CreditCard+Personal.Loan, data = train_data)
ftable(pivot_table)
```
# Question B
To get the answer i.e., the probability of customer accepting the loan offer we need to use the data from the above pivot table. we should divide the number of customers who use online banking, have a credit card and also accepted a loan offer by itself plus number of customers who use online banking and have a credit card but declined the loan offer.

i.e.,
```{r}
probability_acceptloan <- 57/(57+475)*100
print(probability_acceptloan)
```
# Hence, the probability of a customer who holds a credit card using Online banking services accepting Personal loan offer is 10.7%

# Question C
# Create two pivot tables
1. Pivot table with loan as a function of Online Banking
```{r}
ptable_loan.online <- table(train_data$Personal.Loan,
                            train_data$Online)
print(ptable_loan.online)
```
2. Pivot table with loan as a function of Credit Card
```{r}
ptable_loan.CC <- table(train_data$Personal.Loan,
                        train_data$CreditCard)
print(ptable_loan.CC)
```
# Question D
# Probabilities
i.
```{r}
Prob_CCloan <- 91/(91+187)
print(Prob_CCloan)
```
ii.
```{r}
Prob_Onloan <- 179/(179+99)
print(Prob_Onloan)
```
iii. 
```{r}
Prob_loan <- (99+179)/(99+179+1102+1620)
print(Prob_loan)
```
iv.
```{r}
Prob_CCnoloan <- 792/(792+1930)
print(Prob_CCnoloan)
```
v.
```{r}
Prob_Onnoloan <- 1620/(1620+1102)
print(Prob_Onnoloan)
```
vi.
```{r}
Prob_noloan <- (1102+1620)/(1102+1620+99+179)
print(Prob_noloan)
```
# Question E
# Calculate Naive Bayes Probability
```{r}
nbp_loancconline <- (0.0927*0.3273*0.6439)
print(nbp_loancconline)
```
# Hence, Naive Bayes Probability for customers holding Credit Card using Online Banking to accept a personal loan is 0.0195 or 1.95%.

# Question F
# Comparison of Probability calculated in Question B i.e., using pivot table with the Naive Bayes Probability calculated above.

# The Probability calculated using pivot table is 10.71% while it is 1.95% by using Naive Bayes Probability. The probability is 8.76% lower when using Naive Bayes. For the training dataset, pivot table seems accurate but if we are dealing with larger dataset then Naive Bayes might be more accurate.

# Question G

# Calculate Naive Bayes Probability and compare with Question E
```{r}
nbp_loancconline.2 <- naiveBayes(Personal.Loan ~ CreditCard + Online,
                                 data = train_data) 
predict_nbp <- predict(nbp_loancconline.2, 
                       newdata = data.frame( CreditCard = 1,
                       Online = 1), type = "raw")
print(predict_nbp)
```
# Hence, the probability calculated for customers holding Credit Card using Online banking and also accepted loan offer is 0.1157 or 11.57% while the probability calculated in Question E is 0.0195 or 1.95%. The difference is 9.62% which is substantial.















