---
title: "FML Assignment 4"
author: "Jenny Prashanth Buraga"
output: html_document
date: "2023-10-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the dataset and libraries required
```{r}
data <- read.csv("C:/Users/ssim/OneDrive/Desktop/Rhistory/Pharmaceuticals.csv")
library(tidyverse)
library(factoextra)
library(ggplot2)
library(gridExtra)
library(cluster)
```
# Clean the dataset
```{r}
str(data)
data <- na.omit(data)
```

# a. Selecting 9 columns of Numerical variables of the 21 firms in the
dataset which are from column 3 to 11.
```{r}
row.names(data) <- data[,1]
numeric_data <- data[,3:11]
head(numeric_data)
```
# Using scale function to normalize the data
```{r}
scaled_numeric <- scale(numeric_data)
head(scaled_numeric)
```
# Let us perform K-means clustering using multiple values of k and examine the difference in results.
```{r}
kmeans_1 <- kmeans(scaled_numeric, centers = 2, nstart = 30)
kmeans_2 <- kmeans(scaled_numeric, centers = 5, nstart = 30)
kmeans_3 <- kmeans(scaled_numeric, centers = 6, nstart = 30)
Plot_1 <- fviz_cluster(kmeans_1, data = scaled_numeric)+ggtitle("k=2")
Plot_2 <- fviz_cluster(kmeans_2, data = scaled_numeric)+ggtitle("k=5")
Plot_3 <- fviz_cluster(kmeans_3, data = scaled_numeric)+ggtitle("k=6")
grid.arrange(Plot_1,Plot_2,Plot_3, nrow=2)
```
# Let us now determine optimal clusters using elbow method
```{r}
elb_dist <- dist(scaled_numeric, method = "euclidean")
fviz_dist(elb_dist)
```
# Now, we have to calculate the within-cluster sum of sqaures (wss) values for each k value. 
```{r}
set.seed(123)
wss <- function(k){kmeans(scaled_numeric,k,nstart = 10)$tot.withinss
}
kvalues <- 1:10
wss_clusters <- map_dbl(kvalues,wss)
plot(kvalues,wss_clusters, type = "b", pch = 16, frame = T,
     xlab = "Number of clusters",
     ylab = "Total WSS")
```
# The location of the bend in the plot indicates the optimum number of
clusters, so k=5. We will use 5 clusters to perform our final analysis and visualize the results.
```{r}
set.seed(123)
final_analysis <- kmeans(scaled_numeric,5,nstart = 25)
print(final_analysis)
```
# Cluster Plot
```{r}
fviz_cluster(final_analysis, data = scaled_numeric) 
```
# Summary and Cluster plot
```{r}
numeric_data %>%
  mutate(cluster=final_analysis$cluster) %>%
  group_by(cluster) %>% summarise_all("mean")
```
```{r}
clusplot(scaled_numeric,final_analysis$cluster, color=T,
         labels = 2, lines = 0)
```

# b. Interpret the clusters wrt the numeric variables used in forming these clusters.
Cluster 1 - AHM,SGP,WYE,BMY,AZN, ABT, NVS, LLY
Cluster 2 - BAY, CHTT, IVX
Cluster 3 - AGN, PHA
Cluster 4 - JNJ, MRK, PFE,GSK
Cluster 5 - WPI, MRX,ELN,AVE

```{r}
cluster_interpretation <- data[,c(12,13,14)] %>% 
  mutate(clusters = final_analysis$cluster) %>%
  arrange(clusters, ascending = T)
print(cluster_interpretation)
```
# c. Pattern wrt to numerical variables 10 to 12.
```{r}
a1 <- ggplot(cluster_interpretation, 
                    mapping = aes(factor(clusters),
                    fill = Median_Recommendation))+
            geom_bar(position = "dodge")+labs(x="Number of clusters")
a2 <- ggplot(cluster_interpretation, 
                    mapping = aes(factor(clusters),
                    fill = Location))+
            geom_bar(position = "dodge")+labs(x="Number of clusters")
a3 <- ggplot(cluster_interpretation, 
                    mapping = aes(factor(clusters),
                    fill = Exchange))+
            geom_bar(position = "dodge")+labs(x="Number of clusters")
grid.arrange(a1,a2,a3)
```
**As per the graph above, the patterns suggest the following recommendations**

**Cluster 1 suggests to Hold.**

**Cluster 2 also suggests mostly to hold.**

**Cluster 3 suggests to both Hold/Moderate Buy.**

**Cluster 4 suggests to both Hold/Moderate Buy.**

**Cluster 5 suggests to both Moderate Buy/Moderate Sell.**


# d. Providing appropriate name for the clusters.

**Cluster 1 = Investment Cluster.**

**Cluster 2 = Tight Cluster.**

**Cluster 3 = Purchase Cluster.**

**Cluster 4 = Scrutinizing Cluster.**

**Cluster 5 = Examination Cluster.**





